---
title: "Project Assignment - AirBnb"
author: "Group D - Zehao Gao, Antonio Arena, Duy Le"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(naniar)
library(lubridate)
library(forcats)
library(e1071)
library(ggplot2)
library(randomForest)
library(caret)
library(fpp3)
library(tidymodels) #might be few issue loading this library
library(mltools)
library(data.table)
library(recipes)

set.seed(123)

# Setting up Keras and TensorFlow
library(keras)
library(tensorflow)
tensorflow::set_random_seed(123)


```

# Data preprocessing
```{r}
calendar.gz <- read_csv("calendar.csv.gz")
df <- calendar.gz %>%
  mutate(date = as_date(date),
         price = parse_number(price))
df_clean <- df%>% 
  group_by(date) %>% 
  summarise(average_price= mean(price))
df_clean$weekday <- weekdays(df_clean$date)
```

## Create time variables
```{r}

# Create a "weekend" indicator variable
df_clean <- df_clean %>%
  mutate(
    Monday = ifelse(weekday == "Monday", 1, 0),
    Tuesday = ifelse(weekday == "Tuesday", 1, 0),
    Wednesday = ifelse(weekday == "Wednesday", 1, 0),
    Thursday = ifelse(weekday == "Thursday", 1, 0),
    Friday = ifelse(weekday == "Friday", 1, 0),
    Saturday = ifelse(weekday == "Saturday", 1, 0),
    Sunday = ifelse(weekday == "Sunday", 1, 0)
  )
df_clean$month <- months(df_clean$date)
# Create a "Spring indicator variable
df_clean$Spring <- ifelse(df_clean$month %in% c("September","October","November"), 1, 0)
# Create a "Summer" indicator variable
df_clean$Summer <- ifelse(df_clean$month %in% c("December","January","February"), 1, 0)
# Create a "Autumn" indicator variable
df_clean$Autumn <- ifelse(df_clean$month %in% c("March","April","May"), 1, 0)

# Create a "Winter" indicator variable
df_clean$Winter <- ifelse(df_clean$month %in% c("June","July","August"), 1, 0)

# get rid of month and weekday variable
df_clean<- df_clean %>% 
  select(-c(month,weekday))

# Create a "holiday" indicator variable 
df_clean$holiday <- ifelse(df_clean$date %in%
                             as.Date(c("2023-06-12","2023-11-07",
                                       "2023-12-25","2024-01-01",
                                       "2023-12-26","2024-01-26",
                                       "2024-03-11","2024-06-10")), 1, 0)
```

## Split the data into train and test set
```{r}

# Determine the total number of rows in the data
total_rows <- nrow(df_clean)

# Set the number of days to keep for traning, the proportion is 7/10
train_days <- 256

# Create the training set with 256 days
df_train <- df_clean[1:train_days, ]

# Create the testing set with the remaining 110 days
df_test <- df_clean[(train_days + 1):total_rows, ]

df_clean %>%
  as_tsibble() %>%
  gg_season(period = "week") +
  theme(legend.position = "none")

```

## Create dataset for forecasting 295 days to 2024-04-01
```{r}
df_clean <- df_clean %>% as.tibble()
# Create a data frame for forecasting with future dates to 2024-04-01
forecast_dates <- seq(as.Date("2024-06-11"), by = "days", length.out = 295) 
forecast_data <- data.frame(
  date = forecast_dates,
  weekday = weekdays(forecast_dates), 
  month = months(forecast_dates),     
  holiday = rep(0, 295)                
)

forecast_data <- forecast_data %>%
  mutate(
    Monday = ifelse(weekday == "Monday", 1, 0),
    Tuesday = ifelse(weekday == "Tuesday", 1, 0),
    Wednesday = ifelse(weekday == "Wednesday", 1, 0),
    Thursday = ifelse(weekday == "Thursday", 1, 0),
    Friday = ifelse(weekday == "Friday", 1, 0),
    Saturday = ifelse(weekday == "Saturday", 1, 0),
    Sunday = ifelse(weekday == "Sunday", 1, 0)
  )
# Create a "Spring indicator variable
forecast_data$Spring <- ifelse(forecast_data$month %in% c("September","October","November"), 1, 0)
# Create a "Summer" indicator variable
forecast_data$Summer <- ifelse(forecast_data$month %in% c("December","January","February"), 1, 0)
# Create a "Autumn" indicator variable
forecast_data$Autumn <- ifelse(forecast_data$month %in% c("March","April","May"), 1, 0)

# Create a "Winter" indicator variable
forecast_data$Winter <- ifelse(forecast_data$month %in% c("June","July","August"), 1, 0)

# get rid of month and weekday variable
forecast_data<- forecast_data %>% 
  select(-c(month,weekday))

forecast_data$holiday <- ifelse(forecast_data$date %in%
                             as.Date(c("2024-11-05",
                                       "2024-12-25","2024-12-26",
                                       "2025-01-01","2025-01-26",
                                       "2025-03-10")), 1, 0)
```


# Random Forest

##Tune parameters: grid search on mtry
```{r}
# Create control function for training with 10 folds and keep 3 folds for training. search method is grid.
control <- trainControl(method="repeatedcv", number=10, repeats=3, search="grid")
# Define the hyper-parameter grid for mtry values to explore
tunegrid <- expand.grid(.mtry=c(1:12))
# Perform grid search on random forest model, metric is RMSE
rf_gridsearch <- caret::train(average_price~. -date, 
                       data=df_clean, 
                       method="rf", 
                       metric="RMSE", 
                       tuneGrid=tunegrid, 
                       trControl=control)
print(rf_gridsearch)
plot(rf_gridsearch)
# best mtry is 3 with rmse of 41.52707
best_mtry = 3

```

## Grid search on ntrees
```{r}
# Manual search by creat 10 folds and repeat 3 times
control <- trainControl(method = 'repeatedcv',
                        number = 10,
                        repeats = 3,
                        search = 'grid')
# Keep mtry as 4
tunegrid <- expand.grid(.mtry = 3)
modellist <- list()

#train with different ntree parameters
for (ntree in seq(100,1500,100)){
  set.seed(123)
  fit <- caret::train(average_price~.-date,
               data = df_clean,
               method = 'rf',
               metric = 'RMSE',
               tuneGrid = tunegrid,
               trControl = control,
               ntree = ntree)
  key <- toString(ntree)
  modellist[[key]] <- fit
}

#Compare results
results <- resamples(modellist)
summary(results)
#ntree = 1000 for lowest rmse of 41.49274
best_ntree =1000
# Create a graph that shows ntrees and rmse
ntrees <- seq(100, 1500, by = 100)
rmse <- c(41.57047, 41.52107, 41.49400, 41.51544, 41.51058, 41.52255, 41.51234, 41.49305, 41.50383, 41.49274, 41.49472, 41.49350, 41.49869, 41.50550, 41.51121)
tree_rmse <- data.frame(ntrees, rmse)
ggplot(tree_rmse, aes(x = ntrees, y = rmse)) +
  geom_line(color = "#3498db") +
  geom_point(color = "black") +
  labs(title = "RMSE vs. ntrees",
       x = "ntrees",
       y = "RMSE (Root Mean Square Error)") +
  theme_minimal()
```


## Create random forest model, train and check RMSE on test set
```{r}
# Define the Random Forest model
rf_model <- randomForest(average_price ~ .-date, 
                         data = df_train,
                         mtry=best_mtry, 
                         ntree=best_ntree,
                         importance = TRUE)

# Predict using the Random Forest model
test_result_values <- predict(rf_model, newdata = df_test)

test_dates <- seq(as.Date("2024-02-22"), by = "days", length.out = 110)
test_result_df <- data.frame(
  date = as.Date(test_dates),  
  forecastedValue = test_result_values,
  actualPrice = df_test$average_price
)

# show the prediction results together with actual value
colors <- c("Actual" = "#3498db", "Prediction" = "purple")
ggplot(data = test_result_df, aes(x = date)) +
  geom_line(aes(y = actualPrice, color = "Actual"), linetype = "solid") +
  geom_line(aes(y = forecastedValue, color = "Prediction"), linetype = 2) +
  labs(title = "Prediction on test set of Average Price", x = "Date", y = "Average Price") +
  scale_color_manual(name = "Legend", values = colors) +
  scale_linetype_manual(name = "Legend", values = c("Actual" = "solid", "Prediction" = "dashed"))

rmse <- sqrt(mean((df_test$average_price - test_result_values)^2))
cat("Root Mean Squared Error (RMSE) on test set:", rmse, "\n")

```


## Make prediction about 295 days in the futrue which covers the summer period to 2024-04-01
```{r}
# use whole data set to fit the random forest mdel again
rf_model <- randomForest(average_price ~ .-date, 
                         data = df_clean,
                         mtry=best_mtry, 
                         ntree=best_ntree,
                         importance = TRUE)
# Predict using the Random Forest model
forecasted_values <- predict(rf_model, newdata = forecast_data)

# Combine forecasted values and dates into a data frame
forecast_df <- data.frame(
  date = as.Date(forecast_dates),  
  forecastedValue = forecasted_values
)
# Plot the forecasted values
colors <- c("Historical value" = "#3498db", "Forecast value" = "purple")
ggplot() +
  geom_line(data = df_clean, aes(x = date, y = average_price, color = "Historical value"), linetype = "solid") +
  geom_line(data = forecast_df, aes(x = date, y = forecastedValue, color = "Forecast value"), linetype = "solid") +
  labs(title = "Forecasted Average Price for next 295 days", x = "Date", y = "Average Price") +
  scale_color_manual(name = "Legend", values = colors) +
  scale_linetype_manual(name = "Legend", values = c("Historical value" = "solid", "Forecast value" = "solid"))

```

# DNN Model

## Data preprocess and initial model building
```{r}
# Get data into correct form
y_train <- df_train$average_price
x_train <- normalize(as.matrix(df_train %>% select(-c(average_price, date))))

# Initialise list
model_list <- list()
```


```{r}
n_epochs <- 100

# Build model with Dropout
model_list[["Basic"]]$model <- 
  keras_model_sequential() %>%
  layer_dense(units = 128, activation = "relu", input_shape = c(ncol(x_train)),
              kernel_initializer=initializer_random_uniform(minval = -0.05, maxval = 0.05, seed = 123)) %>%
  layer_dropout(rate = 0.4) %>%
  layer_dense(units = 64, activation = "relu",
              kernel_initializer=initializer_random_uniform(minval = -0.05, maxval = 0.05, seed = 123)) %>%
  layer_dropout(rate = 0.3) %>%
  layer_dense(units = 1, activation = "linear",
              kernel_initializer=initializer_random_uniform(minval = -0.05, maxval = 0.05, seed = 123)) %>%
  compile(loss = "mean_squared_error", 
          optimizer = optimizer_rmsprop(), 
          metrics = c("mae")) 

# Train the model
model_list[["Basic"]]$fit <- model_list[["Basic"]]$model %>%
  fit(x_train, 
      y_train, 
      epochs = n_epochs, 
      batch_size = 16, 
      validation_split = 0.2,
       callbacks = callback_early_stopping(monitor = 'val_loss', 
                                          patience = 2))
```


## Get RMSE and Plot from Test Set
```{r}
# Predict using the NN
x_forecast <- as.matrix(df_test %>% select(-c(date, average_price)))

# Predict using the Random Forest model
test_result_values <- predict(model_list$Basic$model, x = normalize(x_forecast))

test_dates <- seq(as.Date("2024-02-22"), by = "days", length.out = 110)


test_result_df <- data.frame(
  date = as.Date(test_dates),  
  forecastedValue = test_result_values,
  actualPrice = df_test$average_price
)
# show the prediction results together with actual value
colors <- c("Actual" = "#3498db", "Prediction" = "purple")
ggplot(data = test_result_df, aes(x = date)) +
  geom_line(aes(y = actualPrice, color = "Actual"), linetype = "solid") +
  geom_line(aes(y = forecastedValue, color = "Prediction"), linetype = 2) +
  labs(title = "Prediction on test set of Average Price", x = "Date", y = "Average Price") +
  scale_color_manual(name = "Legend", values = colors) +
  scale_linetype_manual(name = "Legend", values = c("Actual" = "solid", "Prediction" = "dashed"))

rmse <- sqrt(mean((df_test$average_price - test_result_values)^2))
cat("Root Mean Squared Error (RMSE) on test set:", rmse, "\n")
```


## Make prediction about to 295 days to 2024-04-01
```{r}
# Predict using the NN
x_forecast <- as.matrix(forecast_data %>% select(-c(date)))

forecasted_values <- predict(model_list$Basic$model, x = normalize(x_forecast))


# Combine forecasted values and dates into a data frame
forecast_df <- data.frame(
  date = as.Date(forecast_dates),  
  forecastedValue = forecasted_values
)
# Plot the forecasted values
colors <- c("Historical value" = "#3498db", "Forecast value" = "purple")
ggplot() +
  geom_line(data = df_clean, aes(x = date, y = average_price, color = "Historical value"), linetype = "solid") +
  geom_line(data = forecast_df, aes(x = date, y = forecastedValue, color = "Forecast value"), linetype = "solid") +
  labs(title = "Forecasted Average Price for next 295 days", x = "Date", y = "Average Price") +
  scale_color_manual(name = "Legend", values = colors) +
  scale_linetype_manual(name = "Legend", values = c("Historical value" = "solid", "Forecast value" = "solid"))

```


## Fit 'Improved' NN
This model attempts to add more depth to the NN
```{r}
n_epochs <- 500


# Build model with Dropout
model_list[["Improved"]]$model <- 
  keras_model_sequential() %>%
  layer_normalization() %>%
  layer_dense(units = 64, activation = "relu", input_shape = c(ncol(x_train)),
              kernel_initializer=initializer_random_uniform(minval = -0.05, maxval = 0.05, seed = 123)) %>%
  layer_dense(units = 64, activation = "relu",
              kernel_initializer=initializer_random_uniform(minval = -0.05, maxval = 0.05, seed = 123)) %>%
  layer_dense(units = 1, activation = "linear",
              kernel_initializer=initializer_random_uniform(minval = -0.05, maxval = 0.05, seed = 123)) %>%
  compile(loss = "mean_squared_error", 
          optimizer = optimizer_rmsprop(0.001), 
          metrics = c("mae")) 

# Train the model
model_list[["Basic"]]$fit <- model_list[["Improved"]]$model %>%
  fit(x_train, 
      y_train, 
      epochs = n_epochs, 
      batch_size = 32, 
      validation_split = 0.2,
      callbacks = callback_early_stopping(monitor = 'val_loss', 
                                          patience = 2))
```


## Get RMSE and Plot from Test Set
```{r}
# Predict using the NN
x_forecast <- as.matrix(df_test %>% select(-c(date, average_price)))

# Predict using the Random Forest model
test_result_values <- predict(model_list$Improved$model, x = normalize(x_forecast))

test_dates <- seq(as.Date("2023-05-27"), by = "days", length.out = 110)

test_result_df <- data.frame(
  date = as.Date(test_dates),  
  forecastedValue = test_result_values,
  actualPrice = df_test$average_price
)
# show the prediction results together with actual value
colors <- c("Actual" = "#3498db", "Prediction" = "purple")
ggplot(data = test_result_df, aes(x = date)) +
  geom_line(aes(y = actualPrice, color = "Actual"), linetype = "solid") +
  geom_line(aes(y = forecastedValue, color = "Prediction"), linetype = 2) +
  labs(title = "Prediction on test set of Average Price", x = "Date", y = "Average Price") +
  scale_color_manual(name = "Legend", values = colors) +
  scale_linetype_manual(name = "Legend", values = c("Actual" = "solid", "Prediction" = "dashed"))

rmse <- sqrt(mean((df_test$average_price - test_result_values)^2))
cat("Root Mean Squared Error (RMSE) on test set:", rmse, "\n")
```


## Make prediction about 295 days in the future which covers the summer period
```{r}
# Predict using the NN
x_forecast <- as.matrix(forecast_data %>% select(-c(date)))

forecasted_values <- predict(model_list$Improved$model, x = normalize(x_forecast))


# Combine forecasted values and dates into a data frame
forecast_df <- data.frame(
  date = as.Date(forecast_dates),  
  forecastedValue = forecasted_values
)
# Plot the forecasted values
colors <- c("Historical value" = "#3498db", "Forecast value" = "purple")
ggplot() +
  geom_line(data = df_clean, aes(x = date, y = average_price, color = "Historical value"), linetype = "solid") +
  geom_line(data = forecast_df, aes(x = date, y = forecastedValue, color = "Forecast value"), linetype = "solid") +
  labs(title = "Forecasted Average Price for next 295 days", x = "Date", y = "Average Price") +
  scale_color_manual(name = "Legend", values = colors) +
  scale_linetype_manual(name = "Legend", values = c("Historical value" = "solid", "Forecast value" = "solid"))

```


## Optimisation
Perform a grid search on a single layer NN.
Code is applied from ETC3555 Lab07.

### Evaluate Splits
```{r}
n_epochs <- 2
# Function to evaluate splits
evaluate_split <- function(units, dropout_rate, split, ...) {
  cat("Fitting: split = ", split$id[[1]], "; units = ", units, "; dropout = ", dropout_rate, "\n", sep = "")
  
  # Set seed
  set.seed(123)
  
  # Clear session when function exits
  on.exit(keras::backend()$clear_session())

  # Define a single layer NN with dropout
  model <- keras_model_sequential() %>%
    layer_dense(units = units, activation = "relu", input_shape = c(ncol(x_train)),
                kernel_initializer=initializer_random_uniform(minval = -0.05, maxval = 0.05, seed = 123)) %>%
    layer_dropout(rate = dropout_rate) %>%
    layer_dense(units = 1, activation = "linear",
                kernel_initializer=initializer_random_uniform(minval = -0.05, maxval = 0.05, seed = 123)) %>%
    compile(loss = "mean_squared_error",
            optimizer = optimizer_rmsprop(),
            metrics = c("mae"))
  
  # The data used for modeling (aka the "analysis" set)
  # Note: need to convert back to matrices for keras
  x_train_split <- analysis(split) %>%
    select(starts_with("X")) %>%
    as.matrix()
  
  y_train_split <- analysis(split) %>%
    select(starts_with("Y")) %>%
    as.matrix()

  # Fit to training (analysis) set
  fit(model, x = x_train_split, y = y_train_split, ...)
  
  # Now obtain the holdout set for evaluation
  x_test_split <- assessment(split) %>%
    select(starts_with("X")) %>%
    as.matrix()
  
  y_test_split <- assessment(split) %>%
    select(starts_with("Y")) %>%
    as.matrix()
  
  # The results
  results <- evaluate(model, x_test_split, y_test_split, verbose = 0)
  results[1]
}
```


### Grid Search
```{r}
# Function to perform a grid search
across_grid <- function(split, grid_df, ...) {
  # Execute grid for this resample
  grid_df$loss <- pmap_dbl(
    grid_df,
    evaluate_split,
    split = split,
    ...  # Extra options for `fit`
  )
  # Assigns an split$id[[1]] to each row in the grid_df data frame
  grid_df$id <- split$id[[1]]
  
  grid_df
}

# Convert matrices to dataframes for easier (and quicker) resampling. Need to be a bit fancy because y is a matrix of values (not just a single response).
x_train_df <- as_data_frame(x_train)
names(x_train_df) <- paste0("X", 1:ncol(x_train_df))
y_train_df <- as_data_frame(y_train)
names(y_train_df) <- paste0("Y", 1:ncol(y_train_df))
cal_train <- dplyr::bind_cols(x_train_df, y_train_df)

# 5-fold cross validation
cv_splits <- rsample::vfold_cv(cal_train, v = 5)

# Create grid
grid_df <- list(units = c(16, 32, 64, 128),
                dropout_rate = c(0, .25, .5, .75)) %>%
  purrr::cross_df()

# Fit model to each fold and hyperparameter combination
if (file.exists("tune_results.RData")) {
  load("tune_results.RData")
} else {
  tune_results <- purrr::map_df(
    cv_splits$splits,
    across_grid,
    grid_df = grid_df,
    batch_size = 128,
    epochs = n_epochs,
    verbose = 1,
    view_metrics = 0
  )
  
  save(tune_results, file = "tune_results.RData")
}
```


### Plot heatmap funciton
```{r}
# Plot heatmap function
plot_heatmap <- function(df) {
  ggplot(df, aes(x = units, y = dropout_rate)) + 
    geom_tile(aes(fill = loss)) +
    geom_tile(aes(colour = optimal, size = optimal), fill = NA) +
    geom_text(aes(label = round(loss, 2)), colour = "white") +
    scale_x_log10(breaks = unique(grid_df$units)) +
    scale_fill_gradient(high = "#132B43", low = "#56B1F7") +
    scale_size_manual("optimal", values=c(0, 1)) + 
    scale_colour_manual("optimal", values = c(NA, "darkorange1")) +
    labs(title = "Grid search for optimal dropout rate and number of units",
         subtitle = "Based on 5-fold cross validation")
}
```


### Check average performance by plotting heatmap
```{r}
# Check average performance for units given in question
tune_results %>% 
  filter(units %in% c(16, 32, 64, 128)) %>% 
  group_by(units, dropout_rate) %>% 
  summarise(loss = mean(loss)) %>% 
  ungroup() %>% 
  mutate(optimal = loss == min(loss)) %>% 
  plot_heatmap()
```


### Fit optimal model
```{r}
optimal_units = 128
optimal_dropout = 0

n_epochs = 500
## RMSProp
model_list[["RMSPROP"]]$model <- keras_model_sequential() %>%
    layer_dense(units = optimal_units, activation = "relu", input_shape = c(ncol(x_train)),
                kernel_initializer=initializer_random_uniform(minval = -0.05, maxval = 0.05, seed = 123)) %>%
    layer_dropout(rate = optimal_dropout) %>%
    layer_dense(units = 1, activation = "linear",
                kernel_initializer=initializer_random_uniform(minval = -0.05, maxval = 0.05, seed = 123)) %>%
    compile(loss = "mean_squared_error",
            optimizer = optimizer_rmsprop(),
            metrics = c("mae"))

# Train the model
model_list[["RMSPROP"]]$fit <- model_list[["RMSPROP"]]$model %>%
  fit(x_train, y_train, epochs = n_epochs, batch_size = 128,
      validation_split = 0.2)
```



# LSTM Model

## Preprocessing the data for LSTM inputs
Function to preprocess data for LSTM and after training unscaling + revert differencing was taken from:
http://rwanjohi.rbind.io/2018/04/05/time-series-forecasting-using-lstm-in-r/


```{r lstm-preproc-func}
lag_transform <- function(x, k = 1){
      lagged =  c(rep(NA, k), x[1:(length(x)-k)])
      DF = as.data.frame(cbind(lagged, x))
      colnames(DF) <- c( paste0('x-', k), 'x')
      DF[is.na(DF)] <- 0
      return(DF)
}


scale_data <- function(train, test, feature_range = c(0, 1)) {
  x = train
  fr_min = feature_range[1]
  fr_max = feature_range[2]
  std_train = ((x - min(x) ) / (max(x) - min(x)  ))
  std_test  = ((test - min(x) ) / (max(x) - min(x)  ))
  
  scaled_train = std_train *(fr_max -fr_min) + fr_min
  scaled_test = std_test *(fr_max -fr_min) + fr_min
  
  return( list(scaled_train = as.vector(scaled_train), scaled_test = as.vector(scaled_test) ,scaler= c(min =min(x), max = max(x))) )
  
}


invert_scaling <- function(scaled, scaler, feature_range = c(0, 1)){
  min = scaler[1]
  max = scaler[2]
  t = length(scaled)
  mins = feature_range[1]
  maxs = feature_range[2]
  inverted_dfs = numeric(t)
  
  for( i in 1:t){
    X = (scaled[i]- mins)/(maxs - mins)
    rawValues = X *(max - min) + min
    inverted_dfs[i] <- rawValues
  }
  return(inverted_dfs)
}

series <- df_clean$average_price
series_diff <- difference(series, differences = 1, default = 0)
df_lagged <- lag_transform(series_diff, 1)
         

train_days = 256
total_rows <- nrow(df_lagged)
# Create the training set with 256 days
train <- df_lagged[1:train_days, ]

# Create the testing set with the remaining 110 days
test <- df_lagged[(train_days + 1):total_rows, ]

scaled <- scale_data(train, test, c(-1,1))


y_train <- as_vector(scaled$scaled_train[2])
x_train <- as_vector(scaled$scaled_train[1])

y_test <- as_vector(scaled$scaled_test[2])
x_test <- as_vector(scaled$scaled_test[1])

dim(x_train) <- c(length(x_train), 1, 1)

X_shape2 = dim(x_train)[2]
X_shape3 = dim(x_train)[3]
batch_size = 1
units = 30

```

## LSTM Model Training

```{r lstm-modeling}

lstm_model <- keras_model_sequential() %>%
  layer_lstm(units, 
       batch_input_shape = c(batch_size, X_shape2, X_shape3),
       return_sequences = TRUE,
       stateful = TRUE,
       kernel_initializer=initializer_random_uniform(minval = -0.05, maxval = 0.05, seed = 123)) %>%
  layer_dropout(rate = 0.4) %>%
  layer_lstm(units,
        return_sequences = TRUE,
        stateful = TRUE,
        kernel_initializer=initializer_random_uniform(minval = -0.05, maxval = 0.05, seed = 123)) %>%
  layer_dropout(rate = 0.3) %>%
  layer_dense(units = 1,
              kernel_initializer=initializer_random_uniform(minval = -0.05, maxval = 0.05, seed = 123)) %>%
  compile(loss = 'mean_squared_error',
          optimizer = optimizer_adam(learning_rate = 0.02, weight_decay = 1e-6),
          metrics = c('mae'))

e = 50   
for(i in 1:e ) {
  lstm_model %>% fit(x_train, y_train, epochs=1, batch_size=batch_size, verbose=1, shuffle=FALSE)
  lstm_model %>% reset_states()
}

L = length(x_test)
scaler = scaled$scaler
predictions = numeric(L)

for(i in 1:L){
     X = x_test[i]
     dim(X) = c(1,1,1)
     yhat = lstm_model %>% predict(X, batch_size = batch_size)
     # invert scaling
     yhat = invert_scaling(yhat, scaler,  c(-1, 1))
     # invert differencing
     yhat  = yhat + series[(train_days + i)]
     # store
     predictions[i] <- yhat
}


days <- seq(1, length(series))
# Create a data frame with original data and LSTM predictions
data_df <- data.frame(Date = df_clean$date, OriginalData = series, Prediction = c(rep(NA, train_days), predictions))

# Create the plot 
ggplot(data_df, aes(x = Date)) +
  geom_line(aes(y = OriginalData), color = 'blue') +
  geom_line(aes(y = Prediction), color = 'red') +
  labs(x = 'Date', y = 'Average Price', title = 'LSTM Predictions vs. Original Data') +
  scale_color_manual(values = c('Original Data' = 'blue', 'LSTM Predictions' = 'red')) +
  theme_minimal()

lstm_rmse <- sqrt(mean((df_test$average_price - predictions)^2))
cat("Root Mean Squared Error (RMSE) on test set:", lstm_rmse, "\n")
```

## Forecast using LSTM

```{r lstm-forecast}
x_tr <- as_vector(scaled$scaled_train[1])
x_te <- as_vector(scaled$scaled_test[1])
new_x <- c(x_tr,x_te)

L = nrow(forecast_data)
scaler = scaled$scaler
n_predictions = numeric(L)


for(i in 1:L){
     X = new_x[i]
     dim(X) = c(1,1,1)
     yhat = lstm_model %>% predict(X, batch_size = batch_size)
     # invert scaling
     yhat = invert_scaling(yhat, scaler,  c(-1, 1))
     # invert differencing
     yhat  = yhat + series[i+1]
     # store
     n_predictions[i] <- yhat
}

df_clean <- df_clean %>% as.tibble()
# Create a data frame for forecasting with future dates to 2024-04-01
forecast_dates <- seq(as.Date("2024-06-11"), by = "days", length.out = 295)
forecast_data <- data.frame(
  date = forecast_dates                
)

forecast_df <- data.frame(
  date = as.Date(forecast_dates),  
  forecastedValue = n_predictions
)

test_dates <- seq(as.Date("2024-02-22"), by = "days", length.out = 110)
test_result_df <- data.frame(
  date = as.Date(test_dates),  
  forecastedValue = predictions,
  actualPrice = df_test$average_price
)

# show the prediction results together with actual value
colors <- c("Actual" = "#3498db", "Prediction" = "purple")
ggplot(data = test_result_df, aes(x = date)) +
  geom_line(aes(y = actualPrice, color = "Actual"), linetype = "solid") +
  geom_line(aes(y = forecastedValue, color = "Prediction"), linetype = 2) +
  labs(title = "Prediction on test set of Average Price", x = "Date", y = "Average Price") +
  scale_color_manual(name = "Legend", values = colors) +
  scale_linetype_manual(name = "Legend", values = c("Actual" = "solid", "Prediction" = "dashed"))


# Plot the forecasted values
colors <- c("Historical value" = "#3498db", "Forecast value" = "purple")
ggplot() +
  geom_line(data = df_clean, aes(x = date, y = average_price, color = "Historical value"), linetype = "solid") +
  geom_line(data = forecast_df, aes(x = date, y = forecastedValue, color = "Forecast value"), linetype = "solid") +
  labs(title = "Forecasted Average Price for next 295 days", x = "Date", y = "Average Price") +
  scale_color_manual(name = "Legend", values = colors) +
  scale_linetype_manual(name = "Legend", values = c("Historical value" = "solid", "Forecast value" = "solid"))

```






